Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-01-04 paper/zhang-cook-laa-langrene-menendez.Rmd
modify as per R journal author instruction

---
title: "Visual diagnostics for constrained optimisation with application to guided tours"
author:
  - name: H.Sherry Zhang
    affiliation: Monash University 
    address: Department of Econometrics and Business Statistics
    email:  huize.zhang@monash.edu
  - name: Dianne Cook
    affiliation: Monash University 
    address: Department of Econometrics and Business Statistics
    email:  dicook@monash.edu
  - name: Ursula Laa
    affiliation: University of Natural Resources and Life Sciences
    address: Institute of Statistics
    email:  ursula.laa@boku.ac.at  
  - name: Nicolas Langrené
    affiliation: CSIRO Data61 
    address: 34 Village Street, Docklands VIC 3008 Australia
    email: nicolas.langrene@csiro.au
  - name: Patricia Menéndez
    affiliation: Monash University 
    address: Department of Econometrics and Business Statistics
    email:  patricia.menendez@monash.edu

@@

# Optimisation Methods {#optim}

@@

Several R implementations address optimization problems with both general purpose as well as task specifics solvers. The most prominent one within the general solvers is `optim()` in the \CRANpkg{stats} [@stats] package, which provides both gradient-based and derivative-free optimisation functions. Another general solver specialised in non-linear optimisation is \CRANpkg{nloptr} [@nloptr]. Specific solvers for simulated annealing includes `optim(..., method  "SANN")` and package \CRANpkg{GenSA} [@gensa] that deals with more complicated objective functions. For other task specific solvers, readers are recommended to visit the relevant sections in CRAN task review on *optimisation and mathematical programming* [@crantaskreviewoptim].

@@

The data structure constructed above meets the tidy data principle [@wickham2014tidy] that requires each observation to form a row and each variable to form a column. With tidy data structure, data wrangling and visualisation can ben significantly simplified by well-developed packages such as \CRANpkg{dplyr} [@dplyr] and \CRANpkg{ggplot2} [@ggplot2].

@@

- `explore_trace_search()`: produces summary plots, as shown in Figure \ref{fig:toy-search}
- `explore_trace_interp()`: produces trace plots for the interpolation points, as shown in Figure \ref{fig:toy-interp}
- `explore_space_pca() `: produces plots of projection basis on the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}. Animated version in Figure \ref{fig:toy-pca-animated} can be turned on via the argument `animate = TRUE`
- `explore_space_tour()`: produces animated tour view on the full space of the projection bases, as shown in Figure \ref{fig:toy-tour}.

- `get_*()` extracts and manipulates certain components from the existing data object.
  - `get_best()`: extracts the best basis found in the data object
  - `get_start()`: extracts the starting basis
  - `get_interp()`: extracts the observations in the interpolation
  - `get_search_count()`: produces the summary table of the number of observation in each iteration
  - `get_basis_matrix()`: flattens all the bases into a matrix

- `bind_*()` incorporates additional information outside the tour optimisation into the data object.
  - `bind_theoretical()`: incorporates the best possible basis to the existing data object with the supply of the index function and original data for producing the index value. 
  - `bind_random()`: generates 1000 points on the high dimensional surface of a sphere and binds it to the existing data object and output as a tibble object. `bind_random_matrix()` binds the points to the basis matrix.

- Color
  - `botanical_palettes`:  a collection of color palettes from Australian native plants. Quantitative palettes include daisy, banksia and cherry and sequential palettes contain fern and acacia.
  - `botanical_pal()`: a color interpolator 
  - `scale_color_botanical()`: a ggplot construction for using botanical palettes.

@@

\clearpage

\bibliography{RJreferences}

>>>>

---
title: "Visual Diagnostics for Constrained Optimisation with Application to Guided Tours"
author:
  - name: H.Sherry Zhang
    affiliation: Monash University 
    address: Department of Econometrics and Business Statistics
    email:  huize.zhang@monash.edu
  - name: Dianne Cook
    affiliation: Monash University 
    address: Department of Econometrics and Business Statistics
    email:  dicook@monash.edu
  - name: Ursula Laa
    affiliation: University of Natural Resources and Life Sciences
    address: Institute of Statistics
    email:  ursula.laa@boku.ac.at  
  - name: Nicolas Langrené
    affiliation: CSIRO Data61 
    address: 34 Village Street, Docklands VIC 3008 Australia
    email: nicolas.langrene@csiro.au
  - name: Patricia Menéndez
    affiliation: Monash University 
    address: Department of Econometrics and Business Statistics
    email:  patricia.menendez@monash.edu

@@

# Optimisation methods {#optim}

@@

Several R implementations address optimization problems with both general purpose as well as task specifics solvers. The most prominent one within the general solvers is \code{optim()} in the \CRANpkg{stats} [@stats] package, which provides both gradient-based and derivative-free optimisation functions. Another general solver specialised in non-linear optimisation is \CRANpkg{nloptr} [@nloptr]. Specific solvers for simulated annealing includes \code{optim(..., method  "SANN")} and package \CRANpkg{GenSA} [@gensa] that deals with more complicated objective functions. For other task specific solvers, readers are recommended to visit the relevant sections in CRAN task review on \ctv{optimisation and mathematical programming} [@crantaskreviewoptim].

@@

The data structure constructed above meets the tidy data principle [@wickham2014tidy] that requires each observation to form a row and each variable to form a column. With tidy data structure, data wrangling and visualisation can be significantly simplified by well-developed packages such as \CRANpkg{dplyr} [@dplyr] and \CRANpkg{ggplot2} [@ggplot2].

@@

- \code{explore\_trace\_search()}: produces summary plots, as shown in Figure \ref{fig:toy-search}
- \code{explore\_trace\_interp()}: produces trace plots for the interpolation points, as shown in Figure \ref{fig:toy-interp}
- \code{explore\_space\_pca()}: produces plots of projection basis on the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}. Animated version in Figure \ref{fig:toy-pca-animated} can be turned on via the argument `animate = TRUE`
- \code{explore\_space\_tour()}: produces animated tour view on the full space of the projection bases, as shown in Figure \ref{fig:toy-tour}.

- \code{get\_*()} extracts and manipulates certain components from the existing data object.
  - \code{get\_best()}: extracts the best basis found in the data object
  - \code{get\_start()}: extracts the starting basis
  - \code{get\_interp()}: extracts the observations in the interpolation
  - \code{get\_search\_count()}: produces the summary table of the number of observation in each iteration
  - \code{get\_basis\_matrix()}: flattens all the bases into a matrix

- \code{bind\_*()} incorporates additional information outside the tour optimisation into the data object.
  - \code{bind\_theoretical()}: incorporates the best possible basis to the existing data object with the supply of the index function and original data for producing the index value. 
  - \code{bind\_random()}: generates 1000 points on the high dimensional surface of a sphere and binds it to the existing data object and output as a tibble object. `bind_random_matrix()` binds the points to the basis matrix.

- Color
  - \code{botanical\_palettes}: a collection of color palettes from Australian native plants. Quantitative palettes include daisy, banksia and cherry and sequential palettes contain fern and acacia.
  - \code{botanical\_pal()}: a color interpolator 
  - \code{scale\_color\_botanical()}: a ggplot construction for using botanical palettes.

@@

\bibliography{zhang-cook-laa-langrene-menendez}

### 2021-01-04 paper/zhang-cook-laa-langrene-menendez.tex
modify as per R journal author instruction

% !TeX root = RJwrapper.tex
\title{Visual diagnostics for constrained optimisation with application to
guided tours}
\author{by H.Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, Patricia Menéndez}

@@

\hypertarget{optim}{%
\section{Optimisation Methods}\label{optim}}

@@

Several R implementations address optimization problems with both
general purpose as well as task specifics solvers. The most prominent
one within the general solvers is \texttt{optim()} in the
\CRANpkg{stats} \citep{stats} package, which provides both
gradient-based and derivative-free optimisation functions. Another
general solver specialised in non-linear optimisation is
\CRANpkg{nloptr} \citep{nloptr}. Specific solvers for simulated
annealing includes \texttt{optim(...,\ method\ \ "SANN")} and package
\CRANpkg{GenSA} \citep{gensa} that deals with more complicated objective
functions. For other task specific solvers, readers are recommended to
visit the relevant sections in CRAN task review on \emph{optimisation
and mathematical programming} \citep{crantaskreviewoptim}.

@@

The data structure constructed above meets the tidy data principle
\citep{wickham2014tidy} that requires each observation to form a row and
each variable to form a column. With tidy data structure, data wrangling
and visualisation can ben significantly simplified by well-developed
packages such as \CRANpkg{dplyr} \citep{dplyr} and \CRANpkg{ggplot2}
\citep{ggplot2}.

@@

\begin{itemize}
\item
  \texttt{explore\_trace\_search()}: produces summary plots, as shown in
  Figure \ref{fig:toy-search}
\item
  \texttt{explore\_trace\_interp()}: produces trace plots for the
  interpolation points, as shown in Figure \ref{fig:toy-interp}
\item
  \texttt{explore\_space\_pca()}: produces plots of projection basis on
  the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}.
  Animated version in Figure \ref{fig:toy-pca-animated} can be turned on
  via the argument \texttt{animate\ =\ TRUE}
\item
  \texttt{explore\_space\_tour()}: produces animated tour view on the
  full space of the projection bases, as shown in Figure
  \ref{fig:toy-tour}.
\item
  \texttt{get\_*()} extracts and manipulates certain components from the
  existing data object.

  \begin{itemize}
  \tightlist
  \item
    \texttt{get\_best()}: extracts the best basis found in the data
    object
  \item
    \texttt{get\_start()}: extracts the starting basis
  \item
    \texttt{get\_interp()}: extracts the observations in the
    interpolation
  \item
    \texttt{get\_search\_count()}: produces the summary table of the
    number of observation in each iteration
  \item
    \texttt{get\_basis\_matrix()}: flattens all the bases into a matrix
  \end{itemize}
\item
  \texttt{bind\_*()} incorporates additional information outside the
  tour optimisation into the data object.

  \begin{itemize}
  \tightlist
  \item
    \texttt{bind\_theoretical()}: incorporates the best possible basis
    to the existing data object with the supply of the index function
    and original data for producing the index value.
  \item
    \texttt{bind\_random()}: generates 1000 points on the high
    dimensional surface of a sphere and binds it to the existing data
    object and output as a tibble object.
    \texttt{bind\_random\_matrix()} binds the points to the basis
    matrix.
  \end{itemize}
\item
  Color

  \begin{itemize}
  \tightlist
  \item
    \texttt{botanical\_palettes}: a collection of color palettes from
    Australian native plants. Quantitative palettes include daisy,
    banksia and cherry and sequential palettes contain fern and acacia.
  \item
    \texttt{botanical\_pal()}: a color interpolator
  \item
    \texttt{scale\_color\_botanical()}: a ggplot construction for using
    botanical palettes.
  \end{itemize}
\end{itemize}

>>>>

% !TeX root = RJwrapper.tex
\title{Visual Diagnostics for Constrained Optimisation with Application to
Guided Tours}
\author{by H.Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, Patricia Menéndez}

@@

\hypertarget{optim}{%
\section{Optimisation methods}\label{optim}}

@@

Several R implementations address optimization problems with both
general purpose as well as task specifics solvers. The most prominent
one within the general solvers is \code{optim()} in the \CRANpkg{stats}
\citep{stats} package, which provides both gradient-based and
derivative-free optimisation functions. Another general solver
specialised in non-linear optimisation is \CRANpkg{nloptr}
\citep{nloptr}. Specific solvers for simulated annealing includes
\code{optim(..., method  "SANN")} and package \CRANpkg{GenSA}
\citep{gensa} that deals with more complicated objective functions. For
other task specific solvers, readers are recommended to visit the
relevant sections in CRAN task review on
\ctv{optimisation and mathematical programming}
\citep{crantaskreviewoptim}.

@@

The data structure constructed above meets the tidy data principle
\citep{wickham2014tidy} that requires each observation to form a row and
each variable to form a column. With tidy data structure, data wrangling
and visualisation can be significantly simplified by well-developed
packages such as \CRANpkg{dplyr} \citep{dplyr} and \CRANpkg{ggplot2}
\citep{ggplot2}.

@@

\begin{itemize}
\item
  \code{explore\_trace\_search()}: produces summary plots, as shown in
  Figure \ref{fig:toy-search}
\item
  \code{explore\_trace\_interp()}: produces trace plots for the
  interpolation points, as shown in Figure \ref{fig:toy-interp}
\item
  \code{explore\_space\_pca()}: produces plots of projection basis on
  the reduced space by PCA, as shown in Figure \ref{fig:toy-pca}.
  Animated version in Figure \ref{fig:toy-pca-animated} can be turned on
  via the argument \texttt{animate\ =\ TRUE}
\item
  \code{explore\_space\_tour()}: produces animated tour view on the full
  space of the projection bases, as shown in Figure \ref{fig:toy-tour}.
\item
  \code{get\_*()} extracts and manipulates certain components from the
  existing data object.

  \begin{itemize}
  \tightlist
  \item
    \code{get\_best()}: extracts the best basis found in the data object
  \item
    \code{get\_start()}: extracts the starting basis
  \item
    \code{get\_interp()}: extracts the observations in the interpolation
  \item
    \code{get\_search\_count()}: produces the summary table of the
    number of observation in each iteration
  \item
    \code{get\_basis\_matrix()}: flattens all the bases into a matrix
  \end{itemize}
\item
  \code{bind\_*()} incorporates additional information outside the tour
  optimisation into the data object.

  \begin{itemize}
  \tightlist
  \item
    \code{bind\_theoretical()}: incorporates the best possible basis to
    the existing data object with the supply of the index function and
    original data for producing the index value.
  \item
    \code{bind\_random()}: generates 1000 points on the high dimensional
    surface of a sphere and binds it to the existing data object and
    output as a tibble object. \texttt{bind\_random\_matrix()} binds the
    points to the basis matrix.
  \end{itemize}
\item
  Color

  \begin{itemize}
  \tightlist
  \item
    \code{botanical\_palettes}: a collection of color palettes from
    Australian native plants. Quantitative palettes include daisy,
    banksia and cherry and sequential palettes contain fern and acacia.
  \item
    \code{botanical\_pal()}: a color interpolator
  \item
    \code{scale\_color\_botanical()}: a ggplot construction for using
    botanical palettes.
  \end{itemize}
\end{itemize}
